Think deep for reef resilience

### 2025-07-22 paper/paper_draft.qmd
Add section to discuss the shortcomings of connectivity implementation
And discuss possible improvements from Vogt-Vincent et al. 2023.

It is a common conclusion in coral reef ecology that a primary facet of reef resilience is
the degree of larval connectivity between reefs and the supply of recruits that a reef
receives following disturbance [@hock2017] **expand citations**. In contrast, our results
suggest that there are few differences between resilient and non-resilient reefs in
connectivity metrics such as the ratio of outgoing to incoming larval connectivity and total
connectivity strength. Furthermore, our results suggest no relationship between connectivity
and the ability of reefs to maintain positive carbonate budgets under future conditions.
There has been contrasting evidence [@bode2018; @hock2017] surrounding whether recent
generation connectivity modelling is able to identify reefs that are important for recovery
processes. A key separation of these investigations is their use of differing larval
connectivity models. Reef depth was not considered in these investigations of reef network
resilience. These investigations were performed using prior generations of larval
connectivity models parameterised for 2008-2013 [@hock2017] and 1996-2002 [@bode2018] time
periods. Using updated 1km resolution larval connectivity across the whole GBR, with a
limited temporal sample of particle flows, our results suggest that larval connectivity
alone is not as important for reef resilience and recovery as environmental refugia from
thermal stress such as depth. Our findings may differ due to the highly uncertain nature of
connectivity methods used **citation**, the way larval flows and juvenile recruitment are
implemented in ADRIA - CoralBlox, and the magnitude of early bleaching events decreasing the
breeding capacity of strongly connected reefs. Further improvements and research in
connectivity modelling are necessary to provide a robust comparison of all factors
influencing reef resilience.

>>>>

It is a common conclusion in coral reef ecology that a primary facet of reef resilience is
the degree of larval connectivity between reefs and the supply of recruits that a reef
receives following disturbance [@hock2017; @mumby2008; @torres2024]. In contrast, our
results suggest that there are few differences between resilient and non-resilient reefs in
connectivity metrics such as the ratio of outgoing to incoming larval connectivity and total
connectivity strength. Furthermore, our results suggest no relationship between connectivity
and the ability of reefs to maintain positive carbonate budgets under future conditions.
There has been contrasting evidence [@bode2018; @hock2017] surrounding whether recent
generation connectivity modelling is able to identify reefs that are important for recovery
processes. A key separation of these investigations is their use of differing larval
connectivity models. Reef depth was not considered in these investigations of reef network
resilience. These investigations were performed using prior generations of larval
connectivity models parameterised for 2008-2013 [@hock2017] and 1996-2002 [@bode2018] time
periods. Using updated 1km resolution larval connectivity across the whole GBR, with a
limited temporal sample of particle flows, our results suggest that larval connectivity
alone is not as important for reef resilience and recovery as environmental refugia from
thermal stress such as depth. Our findings may differ due to the highly uncertain nature of
larval connectivity modelling [@vogt-vincent2023], the way larval flows and juvenile
recruitment are implemented in ADRIA - CoralBlox, and the magnitude of early bleaching
events decreasing the breeding capacity of strongly connected reefs. In our investigation
connectivity is implemented as a time-mean measure of reef connectivity over all available
days. Recent analysis of daily reef larval connectivity data suggests that time-mean
connectivity may not be suitable for capturing the seasonal, ecological time scale,
connectivity patterns between locations due to the high daily variability in connectivity
data [@vogt-vincent2023]. This is supported by the connectivity data used in our
investigation, which have large relative standard deviations for connectivity between reefs
**figure**. By incorporating connectivity using methods that capture ecological-timescale
patterns in reef-reef larval dispersal, such as cumulated implicit connectivity
[@vogt-vincent2023], our investigation may find increased importance of both depth and reef
connectivity for resilience, in line with [@hock2017; @torres2024]. Further improvements and
research in connectivity modelling are necessary to provide a robust comparison of all
factors influencing reef resilience.
